---
title: "`easing`"
author: "Chris Hamm"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: TRUE
    toc_float: TRUE
    toc_depth: 3
    css: ../misc/style.css
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = "center", cache = TRUE)
options(dplyr.width = Inf)
```


```{r preliminaries, echo = FALSE, message = FALSE}
set.seed(923741)

library("tidyverse")
library("gganimate")


```{r irisLabs, echo = FALSE}
iris_plot <- iris %>% 
  ggplot(aes(x = Petal.Width, y= Petal.Length)) +
  geom_point(size = 5, alpha = 0.75, show.legend = FALSE) +
  labs(x = "Petal width", y = "Petal length")
```



# Easing

`gganimate` calculates intermediate ("tween") values to transition among states. You can decide how this process occurs, which is called *easing* and can be adjusted with `ease_aes()`.

`ease_aes('linear')`:

  - **quadratic**: \( x^2 \) 
  - **cubic**: \( x^3 \)
  - **quartic**: \( x^4 \)
  - **quintic**: \( x^5 \)
  - **sine**: \( \sin(x) \)
  - **circular**: \( \frac{\pi}{2} \)
  - **exponential**: \( x^e \)
  - **elastic**: elastic release of energy
  - **back**: pullback and release
  - **bounce**: bouncing ball
    - *methods*
      - **-in** - function applied "as-is"
      - **-out** - function applied in reverse
      - **-in-out** - first half applied "as-is", second half is reversed.
      

## `cubic-in-out`
```{r irisEasing, fig.cap = "cubic-in-out easing"}
irisAnimation <- iris_plot +
  transition_states(Species,
                    transition_length = 2, 
                    state_length = 1) +
  labs(title = "Iris {closest_state}")

irisAnimation +
  ease_aes('cubic-in-out')
```


## `bounce-out`
```{r irisBounceOut, fig.cap = "bounce-out easing"}
irisAnimation +
  ease_aes(y = "bounce-out")
```
