---
title: "`tranistion_time()`"
author: "Chris Hamm"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: TRUE
    toc_float: TRUE
    toc_depth: 3
    css: ../misc/style.css
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = "center", cache = TRUE)
options(dplyr.width = Inf)
```


```{r preliminaries, echo = FALSE}
set.seed(923741)

library("tidyverse")
library("gganimate")
library("kableExtra")
library("gapminder")
```

Load the **gapminder data**

```{r irisView, echo = FALSE}
data(gapminder)

gapminder %>% 
  kable() %>% 
  kable_styling("striped", full_width = FALSE) %>% 
  scroll_box(width = "500px", height = "200px")
```



# `transition_time()`
Let's use with the `gapminder` data and try to recreate the Rosling plots.


```{r gapminderStatic}
gap_plot <- gapminder %>% 
  ggplot(
    aes(x = gdpPercap, y = lifeExp, 
        size = pop, color = country)) +
  geom_point(show.legend = FALSE, alpha = 0.7) +
  scale_color_viridis_d() +
  scale_size(range = c(2, 12)) +
  scale_x_log10() +
  labs(x = "GDP per capita", y = "Life expectency")

gap_plot
```


```{r gapTime}
gap_plot +
transition_time(year) +
  labs(title = "Year: {frame_time}")
```


## `facet_wrap()`
```{r gapWrap}
gap_plot +
facet_wrap(~continent) +
  transition_time(year) + 
  labs(title = "Year: {frame_time}")
```


## `view_follow()`

`view_follow(fixed_x = FALSE, fixed_y = FALSE, exclude_layer = NULL, aspect_ratio = 1)`:

  - `fixed_x`: `
  - `fixed_y`:
  - `exclude`: a layer / indes
  - `aspect_ratio`: overrides any `fixed_`

I personally do not care for this in most situations I have encountered. If you find a good use case please let me know.

```{r gapFollow}
gap_plot + 
  transition_time(year) +
  labs(title = "Year: {frame_time}") +
  view_follow(fixed_y = TRUE)
```


## `shadow_wake()`

`shadow_wake(wake_length, size = TRUE, alpha = TRUE, color = NULL, fill = NULL, falloff = "cubic-in", wrap = TRUE, exclude_layer = NULL, exclude_phase = c("enter", "exit"))`

```{r gapShadow}
gap_plot +
transition_time(year) +
  labs(title = "Year: {frame_time}") +
  shadow_wake(wake_length = 0.1, alpha = TRUE)
```
