---
title: "`transition_states()`"
author: "Chris Hamm"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: TRUE
    toc_float: TRUE
    toc_depth: 3
    css: ../misc/style.css
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = "center", cache = TRUE)
options(dplyr.width = Inf)
```


```{r preliminaries, echo = FALSE, message = FALSE}
set.seed(923741)

library("tidyverse")
library("gganimate")
data(iris) # load the iris data
```


```{r iris_plot, echo = FALSE}
iris_plot <- iris %>% 
  ggplot(aes(x = Petal.Width, y= Petal.Length)) +
  geom_point(size = 5, alpha = 0.75, show.legend = FALSE) +
  labs(x = "Petal width", y = "Petal length")

```

With our static plot in place, let's make it move. 


# Motivation

Goal: We want to create an animated figure which shows the `Petal.Length` and `Petal.Width` by `Species` using the `iris` data set.


# Animated plot

We need to use the `transition_states()` function. `transition_states()` works with discrete data. There are other `transition_` funcions but we'll start here.

At a minimum, you need to tell `transition_states()` which discrete variable you want to tranistion between.

```{r irisTransition, fig.cap = "Figure with smooth animation between `Iris` states"}
iris_anim <- iris_plot + 
  transition_states(Species)

iris_anim
```

If you think that this is a fairly straight forward plot and doens't seem to contain a lot of information, you are correct. Right now we want to get the individual parts of the plot working and we will add on to it. This way, if something breaks the code, we'll know exactly where the bug was introduced. Eventually we will want to distinguish `Species` by color. But adding `color = S[ecies` to the `ggplot` call will make the animation choppy, so we'll do that later and I'll show you a cleaver way to smooth it out . 


# Four arguments
`transition_states()` takes **four** arguments:

  - `states` - variable name that contains discrete data
  - `transition_length` - The relative length of the transition (`default = 1`)
  - `state_length` - the relative length of the pause between states (`default = 1`)
  - `wrap` - Should the animate wrap around? (default = `TRUE`) If `TRUE` the last state will be transitioned into the first.

`transition_states(states, transition_length, state_length, wrap = TRUE)`

These transitions look a bit slow. We have considerable power to fine-tune the speed of the transitions. 



```{r iris_fineTune, fig.cap = "2 second transition between states, 1 second pause at each state, wrap = FALSE"}
iris_anim <-  iris_plot +
  transition_states(Species,
                    transition_length = 2, 
                    state_length = 1,
                    wrap = FALSE)
iris_anim
```

What if we change the code so that `wrap = TRUE`?


```{r iris_fineTune2, fig.cap = "2 second transition between states, 1 second pause at each state, wrap = TRUE"}
iris_anim <-  iris_plot +
  transition_states(Species,
                    transition_length = 2, 
                    state_length = 1,
                    wrap = TRUE)
iris_anim
```

Observe how the data points in the animation cycle back  to the start once all the states have been plotted.

Play around with each arguement and see how they alter your figure. 
